---
title: "GGIS 224 Final Project Data Wrangling and Spatial Operations"
author: "Warren Jodjana"
date: "December 14th 2023"
output:   
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
---



# Environment Setup
```{r, echo=TRUE, message = FALSE}
library(raster)
library(dplyr)
library(sf)
library(terra)
library(tmap)
library(readxl)
library(data.table)
library(spatstat)
```


# Clean & Wrangle Data

## Air Quality Index (AQI)

### Load AQI data using `read_excel`
```{r}
airquality <- read_excel("air-quality-index.xlsx")
glimpse(airquality)
```
### Load tract data using `st_read` 

```{r}
airquality_tracts <- st_read("tracts-aqi.geojson")
data.table(airquality_tracts)
```

### Merge data using geoid
```{r}
airquality <- merge(airquality_tracts, airquality, by.x="geoid10", by.y="geoid", all.x=TRUE)
```

### Subset data
```{r}
airquality <- airquality %>%
  select(-c("statefp10", "commarea_n", "commarea", "notes", "countyfp10"))
```

### Inspect data using `data.table`
```{r}
data.table(airquality)
```

### Save cleaned data
```{r}
write.csv(airquality, "airquality.csv")
```

## Traffic & Vehicular Emissions


### Load traffic data using `fread`
```{r}
traffic <- fread("traffic-old.csv")
glimpse(traffic)
```

### Convert to spatial format
```{r}
traffic <- st_as_sf(traffic, coords = c("Longitude", "Latitude"), crs = 4326)
```

### Inspect data using `data.table`
```{r}
data.table(traffic)
```

### Save cleaned data
```{r}
write.csv(traffic, "traffic.csv")
```


## Electric Vehicles

### Load Alternative Fuel data using `fread`
```{r}
alternativefuel <- fread("alternativefuel.csv")
```

### Convert to spatial format
```{r}
alternativefuel <- st_as_sf(alternativefuel, coords = c("Longitude", "Latitude"), crs=4326)
data.table(alternativefuel)
```

### Subset data
```{r}
electric_vehicles <- alternativefuel %>%
  select(`ID`, `Fuel Type Code`, `Station Name`, `Street Address`, `Location`, `geometry`) %>%
  filter(`Fuel Type Code` == "ELEC") %>%
  mutate(`Street Address` = ifelse(`Station Name` == "Paul Simon Chicago JCC", "3348 S Kedzie Ave", `Street Address`)) %>%
  select(-c("Fuel Type Code"))
```

### Inspect data using `data.table`
```{r}
data.table(electric_vehicles)
```


### Save cleaned data
```{r}
write.csv(electric_vehicles, "electric-vehicles.csv")
```


# Spatial Operations

## Traffic Kernel Density Estimation

```{r}
boundary <- st_read("boundaries.geojson")
```

### Ensure both points and boundaries are the same CRS
```{r}
boundary <- st_transform(boundary, crs = 4326)
```

### Creating a two-dimensional point-pattern object

```{r include=TRUE, warning=FALSE}
coords <- st_coordinates(traffic) # extract coordinates from sf object
bbox <- st_bbox(boundary) # define the observation window (bounding box) using the boundary data
window <- owin(xrange = bbox[c("xmin", "xmax")], yrange = bbox[c("ymin", "ymax")])
traffic.ppp <- ppp(coords[,1], coords[,2], window = window) # Create a ppp (two-dimensional point-pattern) object
```

### Perform Kernel Density Estimation

```{r include=TRUE}
kde <- density(traffic.ppp) # perform kernel density estimation
kde_raster <- raster(kde) # converting KDE back to raster form
kde_raster_masked <- mask(kde_raster, boundary) # clip the KDE in raster form to chicago boundaries
```

### Check bandwith and type of kernel

```{r include=TRUE}
str(kde) 
```

<br>

## Kernel Density Estimation on Traffic in Chicago

```{r include=TRUE, warning=FALSE}
tmap_mode("view")
chikde <- tm_shape(boundary) + tm_borders() +
  tm_shape(kde_raster_masked) +
  tm_raster(title="Traffic Density", style="cont", palette="Blues") +
  tm_scale_bar(position = c("left", "bottom"))

chikde
```

## Proximity Analysis

```{r}
nearest_traffic_index <- st_nearest_feature(electric_vehicles, traffic)
nearest_traffic_distances <- st_distance(electric_vehicles, traffic[nearest_traffic_index, ], by_element = TRUE)


tmap_mode("view")
tm <- tm_shape(traffic) +
  tm_bubbles(size = "ID", col = "red", border.col = "black", alpha = 0.5) + 
  tm_shape(electric_vehicles) +
  tm_symbols(col = "blue", size = 0.1) +
  tm_layout(title = "EV Stations and Nearest Traffic Data Points")

tm
```


## Comparative Analysis

```{r}

tmap_mode("plot")
airquality_summary <- airquality %>% 
  group_by(geoid10) %>% 
  summarize(mean_NO2 = mean(CMAQ_NO2, na.rm = TRUE), 
            mean_PM25 = mean(CMAQ_PM25, na.rm = TRUE), 
            mean_O3 = mean(CMAQ_O3, na.rm = TRUE))



plot_NO2 <- tm_shape(airquality_summary) +
  tm_polygons("mean_NO2", title = "Mean NO2") +
  tm_layout(main.title = "Spatial Distribution of NO2", main.title.size = 1)

plot_PM25 <- tm_shape(airquality_summary) +
  tm_polygons("mean_PM25", title = "Mean PM2.5") +
  tm_layout(main.title = "Spatial Distribution of PM2.5", main.title.size = 1)

plot_O3 <- tm_shape(airquality_summary) +
  tm_polygons("mean_O3", title = "Mean O3") +
  tm_layout(main.title = "Spatial Distribution of O3", main.title.size = 1)


tmap_arrange(plot_NO2, plot_PM25, plot_O3)
```